---
title: "Get started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE,
                      comment  = "#>")
```


```{r setup}
library(snakedist)
library(ggplot2)
```


```{r 'ggplot-theme', echo=FALSE}
# Custom ggplot2 theme ----
custom_theme <- function() {
  theme_light() + 
  theme(plot.title   = element_text(face = "bold", family = "serif", size = 18),
        plot.caption = element_text(face = "italic", family = "serif"),
        axis.title   = element_blank(),
        axis.text    = element_text(family = "serif"))
}
```

## Datasets

`snakedist` provides two datasets:

- `adour_lambert93.gpkg`: a spatial layer (`LINESTRING`) of the French river _L'Adour_.
- `adour_sites_coords.csv`: a `data.frame` containing spatial coordinates of 10 locations sampled along the Adour.

The purpose of the package `snakedist` is to compute the distance between each pair of locations by following the shape of the river.

First, let's import the two datasets:


```{r 'import-adour-river', echo=TRUE}
# Import the spatial layer of Adour river ----
path_to_file <- system.file("extdata", "adour_lambert93.gpkg", 
                            package = "snakedist")
adour_river  <- sf::st_read(path_to_file, quiet = TRUE)
adour_river
```


```{r 'import-adour-sites', echo=TRUE}
# Import sites data ----
path_to_file <- system.file("extdata", "adour_sites_coords.csv", 
                            package = "snakedist")
adour_sites  <- read.csv(path_to_file)
adour_sites
```


The function `distance_along()` requires that both layers are spatial objects. So we need to convert the `data.frame` "adour_sites" into an `sf` object.

```{r 'convert-adour-sites', echo=TRUE}
# Convert data.frame to sf object ----
adour_sites <- sf::st_as_sf(adour_sites, coords = c("longitude", "latitude"), 
                            crs = "epsg:2154")
adour_sites
```


Now, we will import a base map of France to visualize our data.

```{r 'get-basemap', echo=TRUE}
# Import France departments ----
france <- rnaturalearth::ne_states("france", returnclass = "sf")

# Remove overseas territories ----
france <- france[grep("^FR\\.", france$"code_hasc"), ]

# Project layer to RGF93 / Lambert-93 system ----
france <- sf::st_transform(france, sf::st_crs(adour_river))
```


Let's map the data.

```{r 'map-adour-river', echo=TRUE, fig.height=9, fig.width=10, out.width='80%', fig.cap='Figure 1. Location of the French river L\'Adour'}
# Get extent of Adour river ----
adour_extent <- sf::st_bbox(adour_river)

# Add some margin around ----
adour_extent <- adour_extent + c(-10000, -10000, 10000, 10000)

# Map ----
ggplot() +
  geom_sf(data = france, fill = "lightgray", col = "white", size = 0.2) +
  geom_sf(data = adour_river, col = "steelblue") +
  geom_sf(data = sf::st_as_sfc(adour_extent), fill = NA, size = 0.3) +
  labs(caption = "RGF93 / Lambert-93 Projection") +
  custom_theme() +
  geom_text(aes(x = 194015, y = 6453703), label = "Atlantic\nOcean",
            color = "lightgray", fontface = "italic", size = 6, 
            family = "serif") +
  geom_text(aes(x = 482630, y = 6150836), label = "Pyrenees",
            color = "lightgray", fontface = "italic", size = 6, 
            family = "serif")
```



```{r 'map-adour-sites', fig.height=9, fig.width=12, out.width='80%', echo=TRUE, fig.cap='Figure 2. Study area with survey sampling'}
ggplot() +
  geom_sf(data = adour_river, col = "steelblue") +
  geom_sf(data = adour_sites, shape = 19, size = 2) +
  labs(caption = "RGF93 / Lambert-93 Projection") +
  custom_theme() +
  geom_sf_label(data = adour_sites, aes(label = site), nudge_x = 5000)
```


