---
title: "Get started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include=FALSE}
knitr::opts_chunk$set(collapse  = TRUE,
                      comment   = "#>",
                      out.width = "100%",
                      dpi       = 96,
                      fig.align = "center")
```


The purpose of the package `snakedist` is to compute the distance between each pair of locations by following the shape of a curved line (e.g. a river).



# Setup

```{r setup}
# Required packages ----
library(snakedist)
library(ggplot2)
```


```{r 'ggplot-theme', echo=TRUE}
# Custom ggplot2 theme ----
custom_theme <- function() {
  theme_light() + 
  theme(plot.title   = element_text(face = "bold", family = "serif", size = 18),
        plot.caption = element_text(face = "italic", family = "serif"),
        axis.title   = element_blank(),
        axis.text    = element_text(family = "serif"))
}
```

# Datasets

`snakedist` provides two datasets:

- `adour_lambert93`: a spatial line layer (`geopackage`) of the French river _L'Adour_.
- `adour_sites_coords`: a table (`csv`) containing spatial coordinates of 5 locations sampled along the Adour.

\

First, let's import the two datasets:


```{r 'import-adour-river', echo=TRUE}
# Import the spatial layer of Adour river ----
path_to_file <- system.file("extdata", "adour_lambert93.gpkg", 
                            package = "snakedist")
adour_river  <- sf::st_read(path_to_file, quiet = TRUE)
adour_river
```

\

```{r 'import-adour-sites', echo=TRUE}
# Import sites data ----
path_to_file <- system.file("extdata", "adour_sites_coords.csv", 
                            package = "snakedist")
adour_sites  <- read.csv(path_to_file)
adour_sites
```

\

The function `distance_along()` (main function of the package) requires that 
both layers are spatial objects. So we need to convert the `data.frame` 
**adour_sites** into an `sf` object.

```{r 'convert-adour-sites', echo=TRUE}
# Convert data.frame to sf object ----
adour_sites <- sf::st_as_sf(adour_sites, coords = c("longitude", "latitude"), 
                            crs = "epsg:2154")
adour_sites
```

\

Let's map the data.

```{r 'map-adour-sites', fig.height=9, fig.width=12, out.width='80%', echo=TRUE, fig.cap='Figure 1. Study area with survey sampling'}
# Visualize data ----
ggplot() +
  geom_sf(data = adour_river, col = "steelblue") +
  geom_sf(data = adour_sites, shape = 19, size = 2) +
  geom_sf_label(data = adour_sites, aes(label = site), nudge_x = 5000) +
  geom_text(aes(x = 334500, y = 6285000), label = "L'Adour", hjust = 0,
            color = "steelblue", fontface = "bold", size = 6, family = "serif") +
  labs(caption = "RGF93 / Lambert-93 Projection") +
  custom_theme()
```


# Compute distances


The function `distance_along()` computes the distance between pairs of sites by following the shape of a linear structure (in our case the river Adour).
It uses the function `sf::st_line_sample()` to regularly (or randomly, argument `type`) sample points on the line with a given density of points (argument `density`). All sampled points between the two sites will be selected and their cumulative euclidean distance will be computed.


```{r 'compute-distances-2', echo=TRUE, eval=FALSE}
# Compute distance along the river ----
dists <- distance_along(adour_sites, adour_river, density = 0.01, type = "regular")
head(dists, 12)
```


```{r 'compute-distances', echo=FALSE}
# Compute distance along the river ----
dists <- suppressMessages(distance_along(adour_sites, adour_river, density = 0.01, type = "regular", mc.cores = 1))
head(dists, 12)
```

This `data.frame` can be converted to a square matrix with the function `df_to_matrix()`.

```{r 'convert-distances', echo=TRUE}
# Convert to square matrix ----
df_to_matrix(dists)
```

\

We can compare these values with the Euclidean distance.

```{r 'compare-distances', echo=TRUE}
# Compare to Euclidean distance ----
coords <- sf::st_coordinates(adour_sites)
dist(coords, upper = TRUE, diag = TRUE)
```

\

This package is related to the package [`chessboard`](https://frbcesab.github.io/chessboard/).
